---
title: "101A_hw_2_annapiskun"
author: "Anna Piskun"
date: "1/17/2020"
output: html_document
---
### Part 1: 
1) Use the armspan.csv data collected from this class.
You will probably need to clean the data before answering these questions. You should justify your choices and decisions for cleaning.

```{r #clean data}

library(readr)
armspan <- read_csv("armspan.csv")
#View(armspan)

# clean up data by taking really low values (such as 5.4) which were most likely recorded in feet and converting them to inches to standardize all values and remove unnecessary outliers. 
armspan[[2,1]] <- 64
armspan[[2,2]] <- 62
armspan[[8,1]] <- 69
armspan[[8,2]] <- 72
armspan[[49,1]] <- 60
# clean up data by taking really high values (such as 150) which was most likely written in centimeters and replacing them with their proper value in inches.
armspan[[6,1]] <- 59
armspan[[6,2]] <- 55

df <- data.frame(armspan)
df
# remove row 15 which contains 38 as an armspan outlier for a person who is 72 inches tall.
df_clean <- df[c(-15),]
df_clean
```

a) Make a plot showing the relationship between armspan and height. Your plot should have armspan on the y-axis. Interpret the plot by describing the trend, strength, and unusual features, if any.

```{r}
x <- df_clean$height
y <- df_clean$armspan

cor(x,y)

plot(x,y, main = "Armspan vs. Height",
     xlab = "Height", ylab = "Armspan",
     xlim = c(59,73), ylim = c(55, 79))

```
